ICT and Future Productivity:\\Evidence and Theory of a GPT\thanks{Correspondence: Department of Economics, Boston College, 140 Commonwealth Avenue, Chestnut Hill, MA 02467. Email: brianti@bc.edu (Marco Brianti) and gati@bc.edu (Laura Gati).}

### 2018-06-09 Latex/Draft_Summer2018/Summer2018.tex
Writing the introduction

Along \cite{comin2006medium}, some theoretical contributions rationalize endogenous productivity dynamics by adapting features of endogenous growth models into DSGE models. Following \cite{romer1990endogenous}, most of those papers augment final-good production functions with an expanding composite of intermediate goods produced by the R\&D sector in order to allow for an endogenous rate of adoption of new technologies.\footnote{\cite{bianchi2014growth}, \cite{anzoategui2016endogenous}, and \cite{moran2017innovation} use similar techniques to endogenize growth. In particular, \cite{bianchi2014growth} augment a DSGE  model using a quality ladders model in the vein of \cite{grossman1991quality}. Moreover, \cite{anzoategui2016endogenous} and \cite{moran2017innovation}, similarly to \cite{comin2006medium}, use a model of expanding variety in the vein of \cite{romer1990endogenous}.} Consistent with those previous models, other papers attempt to provide empirical evidence of a slowdown in the productivity of the R\&D sector. Specifically, they show that although research effort is keeping rising, the rate of new ideas and discoveries is slowing down.\footnote{\cite{jones2009burden} and \cite{bloom2017ideas} are two important contributions that highlight those facts.}

@@

In order to formally explain which mechanism links current ICT to future TFP, we analyze a 2-sector DSGE model which allows ICT to be the general purpose technology of the whole economy.

@@

In line with theories of ICT as a general-purpose technology, we analyze a two-sector general equilibrium model in order to rigorously rationalize previous results and estimate key parameters via impulse-response matching. We conclude that ICT accumulation is able to enhance productivity through a positive spillover effect which takes into account the overall level of diffusion of ICT capital in the economy.

>>>>

Along with \cite{comin2006medium}, some theoretical contributions rationalize endogenous productivity dynamics by adapting features of endogenous growth models into DSGE models. Following \cite{romer1990endogenous}, most of those papers augment final-good production functions with an expanding composite of intermediate goods produced by the R\&D sector in order to allow for an endogenous rate of adoption of new technologies.\footnote{\cite{bianchi2014growth}, \cite{anzoategui2016endogenous}, and \cite{moran2017innovation} use similar techniques to endogenize growth. In particular, \cite{bianchi2014growth} augment a DSGE  model using a quality ladders model in the vein of \cite{grossman1991quality}. Moreover, \cite{anzoategui2016endogenous} and \cite{moran2017innovation}, similarly to \cite{comin2006medium}, use a model of expanding variety in the vein of \cite{romer1990endogenous}.} Consistent with those previous models, other papers attempt to provide empirical evidence of a slowdown in the productivity of the R\&D sector. Specifically, they show that although research effort is keeping rising, the rate of new ideas and discoveries is slowing down.\footnote{\cite{jones2009burden} and \cite{bloom2017ideas} are two important contributions that highlight those facts.}

@@

In order to formally explain which mechanism links current ICT to future TFP, we analyze a 2-sector DSGE model which allows ICT to be the general purpose technology (hereafter GPT) of the whole economy. Both sectoral production functions are fed with three inputs: (i) labor, supplied by households, (ii) hard capital, produced by the final sector, and (iii) ICT capital, produced by the ICT sector. Along with \cite{basu2007information}, we interpret the nature of ICT as a GPT in a neoclassical essence since we assume exogenous technological changes in the ICT sector. When an ICT technology shock arrives, both sectors accumulate ICT capital since it is easier to produce and cheaper to buy. We then assume that an ICT capital deepening positively affects the production function of both sectors through a spillover effect. This positive diffusion effect is intrinsically related to the nature of ICT capital, that is, improving the quality and the speed of communication among economic agents. However, this quality and speed strongly depends on how many economic agents own this technologies. As a simple example, owning a mobile phone enables to contact someone instantaneously if the person you want to reach is endowed by the same technology. As a result, the effectiveness of ICT capital is intrinsically related to its own diffusion. We rationalize this simple line of thought augmenting the production function of each sector in the economy with a spillover effect driven by the diffusion of ICT capital.

@@

Allora ecco perche ICT cresce TFP with some lags...

Matching parameters and results

related literature

structure of the paper
